Report for data : Label Free (non paired)
========================================================

```{r loaddata, echo=FALSE}
outFolder <- paste(
    c('../..', '/outputs/0.05/dpulex6dDMSOC1/vsn_lpe'),
    collapse='')
tempOutputFull <- paste(
    c(outFolder, '/MaxQuant_LFQ_non-linked_vsnlpe_full.csv'), 
    collapse='')
data_proteins <- read.table(tempOutputFull, stringsAsFactors=FALSE, 
                            row.names=NULL, header=TRUE)

h_norm <- c("LFQ.intensity.DMSO.6d.K.1", "LFQ.intensity.DMSO.7d.L.1", 
            "LFQ.intensity.Tamox.C1.7d.O.1", "LFQ.intensity.Tamox.C1.7d.P.1")

h_peptides <- "Peptides"
h_PEP <- "MaxQuant_PEP"
h_ratio <- "Log2.fold_change."
h_ttpval <- "p.value" 
h_ttFDR <- "p.value_BH_corrected"

# data_result <- MaxQuantOutput(folder=intput_folder)
# data_params <- loadParametersData(data_result)

nb_prot <- nrow(data_proteins)
threshold_pval <- 0.05

suppressPackageStartupMessages(library("vsn"))
suppressPackageStartupMessages(library(googleVis))
```

Report for LFQ data, loaded from file : `r tempOutputFull`

This analysis was performed at the Protein Analysis facility of the University of Lausanne (Switzerland), using the R language (see [reference 1](#Ref1)) on the `r format(Sys.time(), "%d/%m/%Y %H:%M:%S")`.

For any further information, please contact us : wwwpaf@unil.ch


Experimental design
-------------------------

Experiment | Type
---------- | ----------
1 | DMSO (Control)
2 | DMSO (Control)
3 | C1 (Concentration 1)
4 | C1 (Concentration 1)

MaxQuant version 1.3.0.5 was used to compute intensities for each experiment/replicate, with database 'C:\MaxQuant\New-fasta\uniprot_Dpulex_20120613.fasta'.
See [reference 2](#Ref2).

Preliminary filter
--------------------------

Each reported intensity used to calculate a fold change should have been computed from at least one peptide (count). 
Proteins should appear in all replicates/channels.


Ratio distribution
-------------------------

```{r citationVSN, echo=FALSE, warning=FALSE}
vsn_citation <- citation("vsn")
vsn_description <- packageDescription("vsn")
```
Normalization was achieved using the Variance Stabilizing method (R package version `r vsn_description$Version`). 
It applies a generalized log2 (glog2) transformation on intensities. See [reference 3](#Ref3).


Output of normalization

```{r violinplot, echo=FALSE, fig.width=10, fig.height=7}
suppressPackageStartupMessages(library("ggplot2"))
matrixdata <- as.matrix(data_proteins[,h_norm])

# Stack data by pasting column after column
data.stacked <- c(matrixdata)

row.names <- row.names(data_proteins)
col.names <- h_norm
numprots <- length(row.names)
num.cond <- 2
num.replicates <- ncol(matrixdata)/num.cond
numallexp <- num.cond * num.replicates

intensitiesdf <- data.frame(
    Intensities = data.stacked,
    Protein.groups = factor(rep(row.names, numallexp)),
    Reporters = factor( rep(col.names, rep(c(numprots), numallexp)) ),
    Conditions = factor( rep(c("Control", "Exp"), c(numprots*num.replicates, numprots*num.replicates)) )
    )

p <- ggplot(data=intensitiesdf, aes(x=Reporters, y=Intensities)) 
p + geom_violin(scale = "count", mapping=aes(fill = Conditions, ylab='glog2( Intensities )'), trim=TRUE, alpha=.75) +
    stat_summary(aes(group=Reporters), fun.y=mean, fun.ymin=min, fun.ymax=max, fill="red", shape=21, size=1) +
    ylab('glog2( Intensities )') +
    xlab('') +
    theme(legend.position="bottom")


```
Clustering of the replicates / conditions after normalization

```{r clusterplot, echo=FALSE, fig.width=8, fig.height=8}
suppressPackageStartupMessages(library(latticeExtra))
#NB: dist computes distances between rows, so we have to transpose data to compute distances between samples
distances <- as.matrix(dist(t(data_proteins[,h_norm]), method="euclidean"))

distances.row <- as.dendrogram(hclust(as.dist(distances)))
row.ordered <- order.dendrogram(distances.row)
legend <- list(
    top=list(
        fun=dendrogramGrob,
        args=list(x=distances.row, side="top"))
    )
levelplot(distances[row.ordered, row.ordered],
          scales=list(x=list(rot=90)), 
          xlab='',
          ylab='',
          legend=legend,
          main="Heatmap")

```

Fold change was computed as follow : 
> Fc = median(3:C1,4:C1)/median(1:DMSO,2:DMSO)

```{r echo=FALSE, warning=FALSE}
#$Ratio=\frac{median(113,114,115,116)}{median(117,118,119,121)}$ 
```

Plot of fold change as function of number of peptides (decreasing).

```{r ratiosplot, echo=FALSE, fig.width=14, fig.height=10}
order_peptides <- order(data_proteins[,h_peptides], decreasing=TRUE)
data_proteins <- data_proteins[order_peptides, ]
is_tt_sig <- as.numeric(data_proteins[,h_ttpval]) < threshold_pval
is_BH_sig <- as.numeric(data_proteins[,h_ttFDR]) < threshold_pval
range_y <- range(data_proteins[,h_ratio])

plot(c(1:nb_prot), data_proteins[,h_ratio], 
     main="Fold changes",
     xlab="Index Peptides (decreasing)", 
     ylab="Log2(Fold change)",
     pch=16,
     col="gray")
grid(col="lightgrey")
abline(h=0)
abline(h=median(data_proteins[,h_ratio]), col="blue", lty=2)

points(c(1:nb_prot)[is_tt_sig], data_proteins[is_tt_sig,h_ratio], pch=16, col="darkorange")
text(c(1:nb_prot)[is_tt_sig], data_proteins[is_tt_sig,h_ratio], data_proteins[is_tt_sig,1], cex=0.7, pos=4, col="darkorange")
if(length(is_BH_sig[is_BH_sig == TRUE])>0) {
    points(c(1:nb_prot)[is_BH_sig], data_proteins[is_BH_sig,h_ratio], pch=16, col="darkred")
    text(c(1:nb_prot)[is_BH_sig], data_proteins[is_BH_sig,h_ratio], data_proteins[is_BH_sig,1], cex=0.7, pos=4, col="darkred")
}

legend(0, range_y[2], 
       legend=c("Non significant", "LPE significant", "LPE&BH significant", "Median of Fold changes"), 
       col=c("gray", "darkorange", "darkred", "blue"),
       fill=c("gray", "darkorange", "darkred", 0),
       border=c("gray", "gray", "gray", 0),
       lty = c(0, 0, 0, 2),
       merge = TRUE)
```


--------------------------------------------------------------------------------------------------------

Significantly different proteins
--------------------------------------------------------------------------------------------------------

```{r citationLPE, echo=FALSE, warning=FALSE}
lpe_citation <- citation("LPE")
```
Data analysis was performed using Local-Pooled-Error method (`r lpe_citation$note` - see [reference 4](#Ref4).
Benjamini-Hochberg method was chosen for multiple testing correction (see [reference 5](#Ref5)).

```{r prepplotT, echo=FALSE}
# Without Benjamini-Hochberg correction
order_pep <- order(data_proteins[,h_PEP], decreasing=FALSE)
data_proteins <- data_proteins[order_pep, ]
is_tt_sig <- as.numeric(data_proteins[,h_ttpval]) < threshold_pval
is_BH_sig <- as.numeric(data_proteins[,h_ttFDR]) < threshold_pval
```

```{r volcanoplotT2, echo=FALSE, fig.width=14, fig.height=10}

pvals <- data_proteins[,h_ttpval]
pvals.correct <- data_proteins[,h_ttFDR]
foldchange <- data_proteins[,h_ratio]
ids <- data_proteins[,1]

plot( foldchange, -log2(pvals),
      main="Volcano plot", sub=paste("(threshold p-value = ",threshold_pval,")"),
      xlab="log2( Fold change )", 
      ylab="-log2( p-value )",
      col="gray", 
      pch=16, cex.lab = 1, cex.axis = 1, cex.main = 1)
grid(col="lightgray")
sd_foldchange <- sd(foldchange)
abline(v=c(c(-2,-1,0,1,2)*sd_foldchange), col="gray", lty=2)

is_test_sig <- as.numeric(pvals) < threshold_pval
abline( h=-log2(threshold_pval), col="orange", lty=2)
points( foldchange[is_test_sig], -log2(pvals[is_test_sig]), pch=16, col="darkorange")
# text( foldchange[is_test_sig], -log2(pvals[is_test_sig]), rownames(matrixdata)[is_test_sig], cex=0.7, pos=4, col="darkorange")

is_corrected_sig <- pvals.correct < threshold_pval
if(length( which(is_corrected_sig == TRUE) ) >0) {
    order.pvals <- order(pvals, decreasing=FALSE)
    reordered.foldchange <- foldchange[order.pvals]
    reordered.pvals <- pvals[order.pvals]
    reordered.pvals.correct <- pvals.correct[order.pvals]
    reordered.ids <- ids[order.pvals]
    is_corrected_sig <- reordered.pvals.correct < threshold_pval
    
    max.pvalcorrect <- max(reordered.pvals.correct[is_corrected_sig])
    ind.maxpval <- which(reordered.pvals.correct[is_corrected_sig]==max.pvalcorrect,arr.ind=TRUE)
    abline( h=-log2( max(reordered.pvals[ ind.maxpval ]) ), col="red", lty=2)
    points( reordered.foldchange[is_corrected_sig], -log2(reordered.pvals[is_corrected_sig]), pch=16, col="darkred")
    text( reordered.foldchange[is_corrected_sig], -log2(reordered.pvals[is_corrected_sig]), reordered.ids[is_corrected_sig], cex=0.7, pos=4, col="darkred")
}

legend("top", 
       legend=c("Non significant", "Non corrected", "Significant after correction", paste("Non-corrected threshold"), paste("Corrected threshold (FDR=",threshold_pval,")")), 
       col=c("gray", "darkorange", "darkred", "orange", "red"),
       fill=c("gray", "darkorange", "darkred", 0, 0),
       border=c("gray", "gray", "gray", 0, 0),
       lty = c(0, 0, 0, 2, 2),
       merge = TRUE,
       cex = 0.8)
```

--------------------------------------------------------------------------------------

Without multiple testing correction, <b>`r length(which(is_tt_sig == TRUE))`</b> significant protein groups were found, with a p-value inferior to `r threshold_pval`.

After Benjamini-Hochberg correction of the p-values, <b>`r length(which(is_BH_sig == TRUE))`</b> significant protein groups were found, on a total of `r length(as.numeric(data_proteins[,1]))`, with an FDR of `r threshold_pval`.


> Please note that, even if some proteins appear as significant with 1 peptide count, most of the scientific journals require a minimum of 2 peptides per identified protein.


--------------------------------------------------------------------------------------

```{r ttesttable, echo=FALSE, results='asis'}
RatiosTable <- gvisTable(data.frame(IDs=ids[is_tt_sig],
    data_proteins[is_tt_sig, 
                  c("UniProt_IDs", "Fasta_headers", "MaxQuant_PEP", "Peptides", 
                    "Log2.fold_change.", "p.value_BH_corrected",
                    "Outlier_Ctrl", "Outlier_Exp")]), 
    options = list(width = 1400, height = 1500))

print(RatiosTable, "chart")
```

--------------------------------------------------------------------------------------

Description of the 'final' table
--------------------------------------------------------------------------------------------------------

Header | Description
------------- | -------------
Protein names | General description of proteins, as found in the UniProtKB database. There can be multiple protein names (= a protein group), as the analysis is sometimes not able to determine uniquely a protein given a set of identified peptides.
UniProt IDs | Identifiers for proteins, as defined in the UniProtKB database. Users can go to the [UniProt website](http://www.uniprot.org/uniprot/) and search for these IDs to have more details on the identified proteins. 
Peptides | Number of peptides used for the identification of a protein group.
MaxQuant PEP | Score of the protein identification, as computed by MaxQuant.
Log2(Fold change) | Relative abundance of the protein.
p-value | Probability of the observations for a given protein, if we would suppose that "there is no change between the conditions". The smaller the p-value, the more confidence we can put in rejecting this supposition.
p-value BH corrected | When a lot of p-values are computed, they have to be corrected for the False Discovery Rate (Type I error). Benjamini-Hochberg correction was used in this analysis. **Proteins whose expression changes significatively should have a corrected p-value between 0 and a decided threshold (`r threshold_pval` in this analysis).** 




***************************************************************************************

> <a id="Ref1">**1.**</a> Please cite the R Project using : `r r_citation <- citation();r_citation$textVersion`
>
> <a id="Ref2">**2.**</a> MaxQuant software : published in MaxQuant enables high peptide identification rates, individualized p.p.b.-range mass accuracies and proteome-wide protein quantification by Jürgen Cox & Matthias Mann (2008)
> 
> <a id="Ref3">**3.**</a> R package for VSN : published in `r vsn_citation$title` by `r vsn_citation$author`
>
> <a id="Ref4">**4.**</a> R package for LPE : published in `r lpe_citation$title` by `r lpe_citation$author`
> 
> <a id="Ref5">**5.**</a> Benjamini-Hochberg correction : published in "Controlling the false discovery rate: a practical and powerful approach to multiple testing" by Yoav Benjamini & Yosef Hochberg (1995)

